## Fig 4: Forecast horizon and Rain Error

Visualize extent to which forecast horizon (Days Out) affects forecast accuracy.

Filter dataset by DO 0 through 6 and aggregate.

```{r}
# Create new dataset for Days Out analysis.
precip_acc_do <- tibble(ProbPrecip_forecast=seq(0,100,10))
j <- 2

for (i in seq(0, 6)){
  weather_export_do <- weather_export %>% 
    filter(Days_out==i) %>% 
    #mutate(Precip = PrecipIn >= 0.01) %>% 
    as.data.table()
  
  pa <- weather_export_do[,j=list(round(mean(Precip, na.rm=T)*100, 2)),
                       by=list(ProbPrecip_forecast)] %>% 
    arrange(ProbPrecip_forecast)
  
  # 6 DO has no forecasts of 100%
  if (i==6){
    pa <- rbind(pa, tibble(ProbPrecip_forecast=100,V1=NA))
  }
  
  precip_acc_do <- cbind(precip_acc_do, pa[,2])
  
  colnames(precip_acc_do)[j] <- paste("do",i,sep="")
  
  j <- j+1
}

#Be mindful of parenthesis in geom_line with regards to color, as well as scale_colour_manual alphabetizing colors.

ggplot(precip_acc_do) +
  geom_line(aes(x=ProbPrecip_forecast, y=do0,
            color="#d0d1e6"), size=.8) +
  geom_line(aes(x=ProbPrecip_forecast, y=do1,
            color="#a6bddb"), size=.8) +
  geom_line(aes(x=ProbPrecip_forecast, y=do2,
            color="#67a9cf"), size=.8) +
  geom_line(aes(x=ProbPrecip_forecast, y=do3,
            color="#3690c0"), size=.8) +
  geom_line(aes(x=ProbPrecip_forecast, y=do4,
            color="#02818a"), size=.8) +
  geom_line(aes(x=ProbPrecip_forecast, y=do5,
            color="#016c59"), size=.8) +
  geom_line(aes(x=ProbPrecip_forecast, y=do6,
            color="#014636"), size=.8) +
  geom_abline(size=.8, linetype="dashed") +
  theme_light() +
  scale_colour_manual(name = 'Days Out', 
         values =c('#d0d1e6'='#d0d1e6','#a6bddb'='#a6bddb','#67a9cf'='#67a9cf', '#3690c0'='#3690c0','#02818a'='#02818a','#016c59'='#016c59', '#014636'='#014636'), 
         labels = c('6','5','4','3','2','1','0')) +
  scale_x_continuous(breaks = seq(0,100,20)) +
  scale_y_continuous(breaks = seq(0,100,20)) +
  labs(x="Expected Proportion of Rainy Days", y="Actual Proportion of Rainy Days") +
  theme(axis.title.x = element_text(size=17),
        axis.title.y = element_text(size=17),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14))
```

Table of RE by DO

```{r}
mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do0), na.rm=T)

re_table_do <- tibble(
  re0 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do0), na.rm=T),
  re1 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do1), na.rm=T),
  re2 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do2), na.rm=T),
  re3 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do3), na.rm=T),
  re4 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do4), na.rm=T),
  re5 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do5), na.rm=T),
  re6 = mean(abs(precip_acc_do$ProbPrecip_forecast - precip_acc_do$do6), na.rm=T)
)
```



## Fig 5: Scatter of Number of Rainy Days by MARE

Scatter
```{r}
p1 <- ggplot(rain_metrics) +
  geom_point(aes(x=rainy_days_py, y=re), size=2) +
  theme_bw() +
  labs(x="Number of Rainy Days in a Year", y="Mean Absolute Rain Error") +
  theme(axis.title.x = element_text(size=17),
        axis.title.y = element_text(size=17),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14),
        legend.text=element_text(size=14))

p1

#ggsave('pics/mare by nbr rainy days.jpeg', plot=p1,width=7, height=6)
```


## Correlation between Number of Rainy Days and MARE
```{r}
cor(rain_metrics[,c("rainy_days_py","re")])
psych::corr.test(rain_metrics[,c("rainy_days_py","re")])

# cor with top 5 raniest cities removed
rain_metrics <- rain_metrics %>% arrange(desc(rainy_days_py))
cor(rain_metrics[-(1:5),c("rainy_days_py","re")])
psych::corr.test(rain_metrics[-(1:5),c("rainy_days_py","re")])
```

## Fig 6: Least accurate cities (Blue purple graphs)
Comparing rainy days per year and inches of rain per rainy day for highest Rain Error cities.

```{r}
rain_graph <- rain_metrics

# Normalize
rain_graph[,c("rainy_days_py","avg_precip_in")] <- lapply(rain_graph[,c("rainy_days_py","avg_precip_in")],scale)

# Wide to long
rain_graph <- melt(rain_graph[,c("city","re","rainy_days_py","avg_precip_in")], id=c("city", "re"))


# without legend
rhg_cols <- c("#386cb0", "#beaed4")

# Top 10 RE cities. Use n=20 because dataset is long format with 2 vars.
p1 <- ggplot(rain_graph %>% top_n(20,re)) + 
  geom_bar(mapping=aes(x=reorder(city, re), y=value, fill=variable), stat="identity", position="dodge") +
  labs(y="Normalized values", x=NULL) +
  theme_bw() + 
  scale_fill_manual(values = rhg_cols, name=NULL, labels=c("Normalized avg nbr of rainy days per yr","Normalized avg in. of precipitation per rainy day")) +  
  scale_y_continuous(minor_breaks = NULL) +
  theme(axis.title.x = element_text(size=17),
        axis.title.y = element_text(size=17),
        axis.text.x = element_text(size=14, angle=315, hjust=.1),
        axis.text.y = element_text(size=14),
        legend.text=element_text(size=10),
        legend.position='bottom')

p1

#ggsave('C:/Google Drive/CSUEB/data_expo/final analysis/pics/new graphs/most accurate cities bar.jpeg', plot=p1, width=8.5, height=6)
```




## Fig 7: Most accurate cities
Comparing rainy days per year and inches of rain per rainy day for lowest Rain Error cities.

```{r}
# Top 10 most accurate cities (lowest RE).
# without legend
rhg_cols <- c("#386cb0", "#beaed4")

p1 <- ggplot(rain_graph %>% top_n(20,desc(re))) + 
  geom_bar(mapping=aes(x=reorder(city, re), y=value, fill=variable), stat="identity", position="dodge") +
  labs(y="Normalized values", x=NULL) +
  theme_bw() + 
  scale_fill_manual(values = rhg_cols, name=NULL, labels=c("Normalized avg nbr of rainy days per yr","Normalized avg in. of precipitation per rainy day")) +  
  scale_y_continuous(minor_breaks = NULL, limits=c(-2,1)) +
  theme(axis.title.x = element_text(size=17),
        axis.title.y = element_text(size=17),
        axis.text.x = element_text(size=14, angle=315, hjust=0.1),
        axis.text.y = element_text(size=14),
        legend.text=element_text(size=10),
        legend.position = 'bottom')

p1

#ggsave('C:/Google Drive/CSUEB/data_expo/final analysis/pics/new graphs/least accurate cities bar.jpeg', plot=p1, width=8.5, height=6)
```